---
title: "Reproducible Research - Course project 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

1. Code for reading in the dataset and/or processing the data

```{r load}
Activity <- read.csv(file="activity.csv", header=TRUE, sep=",")
head(Activity)
Activity$date<-as.POSIXct(Activity$date, format = "%Y-%m-%d")

```

2. Histogram of the total number of steps taken each day

```{r histogram}
Activity$date<-as.POSIXct(Activity$date, format = "%Y-%m-%d")
ActivityPerDay<- aggregate(steps ~ date, Activity ,sum)
head(ActivityPerDay)
hist(ActivityPerDay$steps)
```

3. Mean and median number of steps taken each day

``` {r MeanMedia}
ActivityMean <- mean(ActivityPerDay$steps)
ActivityMedian <- median(ActivityPerDay$steps)
ActivityMean
ActivityMedian
```

Mean is `r ActivityMean` and median is `r ActivityMedian` .

4. Time series plot of the average number of steps taken

``` {r SeriesPlotAverage}
ActivityPerInterval<- aggregate(steps ~ interval, Activity , mean)
head(ActivityPerInterval)
plot(ActivityPerInterval$interval, ActivityPerInterval$steps, type = 'l', xlab = "5 min interval", ylab = "Average number of steps")
```

5. The 5-minute interval that, on average, contains the maximum number of steps

``` {r MaximunNumber}
ActivityMaximun <- max(ActivityPerInterval$steps)
ActivityIntervalMaximun <- ActivityPerInterval[ActivityPerInterval$steps == ActivityMaximun,]$interval
```

The interval with maximun number of steps (on average) is `r ActivityIntervalMaximun` with `r ActivityMaximun` number of steps.

6. Code to describe and show a strategy for imputing missing data

``` {r MissingData}
sum(is.na(Activity$steps))
sum(is.na(Activity$interval))
sum(is.na(Activity$date))
ActivityNACount <- sum(is.na(Activity$steps))
```

The total number of missing values in the dataset is `r ActivityNACount`.  Only the field steps contains NA values.

The strategy used to fill in the NA values is take the mean for that 5-minute interval.

``` {r CleanActivity}
CleanActivity <- Activity
for (i in 1:dim(Activity)[][1]) {
    interval <- CleanActivity[i,]$interval
    mean5min <- ActivityPerInterval[1,]$steps
    CleanActivity$steps[is.na(CleanActivity$steps)] <- mean5min
 }
head(CleanActivity)
```

7. Histogram of the total number of steps taken each day after missing values are imputed

```{r CleanHistogram}
CleanActivityPerDay<- aggregate(steps ~ date, CleanActivity ,sum)
head(CleanActivityPerDay)
hist(CleanActivityPerDay$steps)
```

Mean and median number of steps taken each day.

``` {r CleanMeanMedia}
CleanActivityMean <- mean(CleanActivityPerDay$steps)
CleanActivityMedian <- median(CleanActivityPerDay$steps)
CleanActivityMean
CleanActivityMedian
```

Mean is `r CleanActivityMean` and median is `r CleanActivityMedian` .

There are a clear impact because the NA value on the mean an the media:  
   -  Mean: with NA values is `r ActivityMean` and without is `r CleanActivityMean`  
   -  Median: with NA value is `r ActivityMedian` and without is `r CleanActivityMedian`  
   
9. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

``` {r}
library(dplyr)
ActiviyWeekday <- Activity
ActivityWeekday <- mutate(ActiviyWeekday, weekDay = weekdays(date))
ActivityWeekday <- mutate(ActiviyWeekday, isWeekend = ifelse(ActivityWeekday$weekDay == "Saturday" |    ActivityWeekday$weekDay == "Sunday" , "weekend", "weekday"))
ActivityWeekdayPerInterval<- aggregate(steps ~ interval + isWeekend, ActivityWeekday , mean)
head(ActivityWeekdayPerInterval)
library(lattice)
xyplot(steps ~ interval | isWeekend, data = ActivityWeekdayPerInterval, type = "l", layout = c(1, 2), xlab = "5 min interval", ylab = "Average number of steps")

```